Categorical Foundations of E-Graphs With Binding

### 2025-07-08 POPL2026/paper/CategoricalFoundationsOfEGraphsWithBinding.tex
notes and popl template
start section 7. lambda calculus w/ explicit substitution

\documentclass[]{IEEEtran}
% \IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%Template version as of 6/27/2024

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor,soul}

@@

\usepackage{mathtools}
\usepackage{tikz}
\usepackage{tikz-network}
\usepackage{xspace}
\usepackage{xargs}
\usepackage{fontenc}

@@

\usepackage[final]{microtype}

@@

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{E-Graphs With Bindings
}

% \author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
% \IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
% \textit{name of organization (of Aff.)}\\
% City, Country \\
% email address or ORCID}
% \and
% }

\author{Anonymous}

@@

\input{sections/introduction}
\input{sections/categorical_semantics}
\input{sections/combinatorial_semantics}
\input{sections/soundness}

@@

\input{sections/appendix}

>>>>

\documentclass[acmsmall,screen, nonacm, review, anonymous, authoryear]{acmart}

\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor,soul}

@@

\usepackage{mathtools}
\usepackage{tikz}
\usepackage{tikz-network}
\usepackage{xspace}
\usepackage{xargs}
\usepackage{fontenc}
\usepackage{csquotes}

@@

\usepackage{microtype}

@@

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\title{Categorical Foundations of E-Graphs With Binding}

\begin{document}

@@

\input{sections/introduction}
\input{sections/categorical_semantics}
\input{sections/combinatorial_semantics}
\input{sections/soundness}
\input{sections/lambda-calculus}

\section{Conclusion}
In this work we introduced e-graphs with bindings formalised as morphisms of a closed $\catname{SLat}$-SMC.
We then provided a combinatorial representation for such morphisms in terms of hierarchical hypergraphs (e-hypergraphs) with a suitable notion of sound and complete DPO rewriting, making it possible for an e-hypergraph to represent equivalences derivable from a set of equations $\mathcal{E}$.
An important future work is exploring whether such combinatorial structures may lead to better algorithmics for doing equality saturation for $\lambda$-calculus.

@@

% \input{sections/appendix}
